---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dev = "ragg_png"
)
```

# nswtheme

<!-- badges: start -->
[![R-CMD-check](https://github.com/digitalnsw/nswtheme/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/digitalnsw/nswtheme/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of `nswtheme` is to provide a R tool kit for styles, patterns and
standards following the [Digital NSW, NSW Design System](https://github.com/digitalnsw/nsw-design-system)


## Installation

You can install nswtheme like so:

``` r
# CRAN release (not yet!)
# install.packages('nswtheme')

# development version
install.packages('pak')
pak::pak('digitalnsw/nswtheme')
```

## Usage
In many instances it may be sufficient to set the global theme:

```{r}
library(ggplot2)
library(nswtheme)

set_theme(theme_nsw())

ggplot(mpg, aes(displ, hwy, colour = class)) +
  geom_point()
```

More control is available through palette functions such as `pal_waratah()`.
See `vignette("nswtheme")` for usage guidelines.
