---
title: "Demonstration gallery"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  dev = "ragg_png",
  collapse = TRUE,
  echo = FALSE,
  fig.width = 3.5,
  fig.height = 3.5,
  comment = "#>"
)
```

```{r setup}
library(nswtheme)
library(ggplot2, warn.conflicts = FALSE)
library(dplyr, warn.conflicts = FALSE)
library(forcats)
library(reactable)
```

```{r, echo = TRUE}
set_theme(
  theme_nsw(variant = "corporate") + theme(legend.position = "bottom")
)
```

```{r}
applications <- tibble::tribble(
  ~`Local government area` , ~`Applications assessed` ,
  "City of Sydney"         ,                      312 ,
  "Parramatta"             ,                      284 ,
  "Blacktown"              ,                      251 ,
  "Liverpool"              ,                      198 ,
  "Penrith"                ,                      176 ,
  "Camden"                 ,                      143
)

support <- tibble::tribble(
  ~Quarter , ~Phone , ~Digital , ~`In person` ,
  "Q1"     ,   4200 ,     5800 ,         1100 ,
  "Q2"     ,   3900 ,     6400 ,          980 ,
  "Q3"     ,   3600 ,     7100 ,          890 ,
  "Q4"     ,   3200 ,     8200 ,          760
)

road_incidents <- tibble::tribble(
  ~Month      , ~Incidents ,
  "January"   ,       3240 ,
  "February"  ,       2980 ,
  "March"     ,       3100 ,
  "April"     ,       2760 ,
  "May"       ,       2640 ,
  "June"      ,       2580 ,
  "July"      ,       2490 ,
  "August"    ,       2520 ,
  "September" ,       2710 ,
  "October"   ,       2890 ,
  "November"  ,       3040 ,
  "December"  ,       3380
)

service_transactions <- tibble::tribble(
  ~Period    , ~Web  , ~App  , ~`Assisted digital` ,
  "Jul 2024" , 42000 , 18000 ,                6200 ,
  "Jan 2025" , 48000 , 24000 ,                5400 ,
  "Jun 2025" , 51000 , 31000 ,                4100
)

grants <- tibble::tribble(
  ~Month      , ~`Cumulative total ($M)` ,
  "July"      , 12.4                     ,
  "August"    , 28.7                     ,
  "September" , 41.2                     ,
  "October"   , 58.9                     ,
  "November"  , 74.3                     ,
  "December"  , 89.1
)

open_cases <- tibble::tribble(
  ~Period , ~Standard , ~High , ~Critical ,
  "Jan"   ,       200 ,   540 ,        80 ,
  "Feb"   ,       230 ,   620 ,        90 ,
  "Mar"   ,       300 ,   640 ,        95 ,
  "Apr"   ,       330 ,   680 ,       120 ,
  "May"   ,       520 ,   690 ,       135 ,
  "Jun"   ,       600 ,   700 ,       150
)

planning <- tibble::tribble(
  ~Application , ~`Complexity score` , ~`Processing time (days)` ,
           "A" ,                  24 ,                         4 ,
           "B" ,                  31 ,                         6 ,
           "C" ,                  45 ,                         9 ,
           "D" ,                  62 ,                        14 ,
           "E" ,                  74 ,                        18 ,
           "F" ,                  86 ,                        22
)
```

Examples derived from the NSW Design System [guidance on charts and graphs](https://designsystem.nsw.gov.au/docs/content/methods/charts-and-graphs.html).
This page is intended to demonstrate how the charts from the guidance page look with minimal changes beyond setting the theme.
There are some inevitable differences due to different defaults in chart behaviours between ggplot and other frameworks.

## Comparison and ranking

<div class="container">
<div class="row">
<div class="col">

### Bar (vertical)

```{r}
applications |>
  mutate(
    `Local government area` = fct_reorder(
      `Local government area`,
      desc(`Applications assessed`)
    )
  ) |>
  ggplot() +
  geom_col(aes(`Local government area`, `Applications assessed`)) +
  scale_x_discrete(guide = guide_axis(angle = 30)) +
  scale_y_continuous(expand = expansion(mult = c(0, .2))) +
  theme(plot.margin = margin_part(l = 20)) +
  labs(
    title = "Development applications assessed",
    subtitle = "By local government area, January to March 2025",
    x = NULL,
    y = NULL
  )
```

**Summary**: City of Sydney and Parramatta assessed the highest number of applications in Q1 2025, together accounting for over a third of the total.

```{r}
reactable(applications, theme = reactable_nswtheme())
```

**Source**: NSW Planning Portal. Last updated: March 2025.

</div>
<div class="col">

### Stacked bar

```{r}
support |>
  tidyr::pivot_longer(-Quarter) |>
  ggplot() +
  geom_col(aes(Quarter, value, fill = name)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  labs(
    title = "Support requests by contact type",
    subtitle = "Volume by channel, 2024-25 financial year",
    fill = NULL,
    x = NULL,
    y = NULL
  )
```

**Summary**: Digital contact grew steadily across all four quarters, while phone and in-person volumes declined. By Q4, digital accounted for more than two-thirds of all requests.

```{r}
reactable(support, theme = reactable_nswtheme())
```

**Source**: Service NSW. Last updated: June 2025.

</div>
</div>
</div>

## Trends over time

<div class="container">
<div class="row">
<div class="col">

### Line

```{r}
road_incidents |>
  mutate(Month = fct_inorder(Month, ordered = TRUE)) |>
  ggplot(aes(Month, Incidents)) +
  geom_line(aes(group = NA)) +
  geom_point() +
  scale_x_discrete(
    labels = \(x) substr(x, 1, 3),
    guide = guide_axis(angle = 45)
  ) +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.05))) +
  labs(
    title = "Road incident reports",
    subtitle = "NSW monthly total, January to December 2024",
    x = NULL,
    y = NULL
  )
```

**Summary**: Incidents were highest in summer months, peaking in December and January. July recorded the lowest monthly total for the year.

```{r}
reactable(road_incidents, theme = reactable_nswtheme())
```

**Source**: Transport for NSW. Last updated: December 2024.

</div>
<div class="col">

### Multi-line

```{r}
service_transactions |>
  mutate(Period = fct_inorder(Period, ordered = TRUE)) |>
  tidyr::pivot_longer(-Period) |>
  ggplot(aes(Period, value)) +
  geom_line(aes(linetype = name, colour = name, group = name)) +
  geom_point(aes(shape = name, colour = name)) +
  scale_y_continuous(
    limits = c(0, NA),
    expand = expansion(mult = c(0, 0.2)),
    labels = scales::label_comma()
  ) +
  labs(
    title = "Online service transactions by channel",
    subtitle = "Monthly volume, July 2024 to June 2025",
    colour = NULL,
    shape = NULL,
    linetype = NULL,
    x = NULL,
    y = NULL
  )
```

**Summary**: App transactions grew fastest over the year, nearly doubling. Web remained the largest channel. Assisted digital declined steadily as self-service adoption increased.

```{r}
reactable(service_transactions, theme = reactable_nswtheme())
```

**Source**: Service NSW. Last updated: June 2025.

</div>
</div>
<div class="row">
<div class="col">

### Area

```{r}
grants |>
  mutate(Month = fct_inorder(Month, ordered = TRUE)) |>
  ggplot(aes(Month, `Cumulative total ($M)`)) +
  geom_area(aes(group = NA, fill = after_scale(alpha(fill, 0.3)))) +
  geom_point() +
  scale_x_discrete(labels = \(x) substr(x, 1, 3)) +
  labs(
    title = "Grant funding disbursed",
    subtitle = "Running total by month, July to December 2024",
    x = NULL,
    y = NULL
  )
```

**Summary**: Grant disbursements accelerated from October, with the final quarter accounting for more than half of the full-year total.

```{r}
reactable(grants, theme = reactable_nswtheme())
```

**Source**: NSW Treasury. Last updated: December 2024.

</div>
<div class="col">

### Stacked area

```{r}
open_cases |>
  mutate(Period = fct_inorder(Period, ordered = TRUE)) |>
  tidyr::pivot_longer(-Period) |>
  ggplot(aes(Period, value)) +
  geom_area(aes(fill = name, group = name)) +
  labs(
    title = "Support cases open by priority level",
    subtitle = "Cumulative open cases, January to June 2025",
    fill = NULL,
    x = NULL,
    y = NULL
  )
```

**Summary**: The total number of open cases grew steadily across all priority levels. Critical cases grew fastest proportionally, nearly doubling between January and June.

```{r}
reactable(open_cases, theme = reactable_nswtheme())
```

**Source**: Service NSW. Last updated: June 2025.

</div>
</div>
</div>


## Distribution and correlation

<div class="container">
<div class="row">
<div class="col">

### Scatter plot

```{r}
planning |>
  ggplot(aes(`Complexity score`, `Processing time (days)`)) +
  geom_point() +
  geom_blank(aes(x = x, y = y), data = data.frame(x = 0, y = 0)) +
  labs(
    title = "Application complexity and processing time",
    subtitle = "Individual applications, Q1 2025",
    x = NULL,
    y = NULL
  )
```

**Summary**: Applications with higher complexity scores consistently took longer to process. Most straightforward applications (score below 30) resolved within 5 working days. Complex applications (score above 70) averaged 18 days, with significant variation.

```{r}
reactable(planning, theme = reactable_nswtheme())
```

**Source**: NSW Planning Portal. Last updated: March 2025.

</div>
<div class="col">

</div>
</div>
</div>
